Physics-based protein structure prediction and design using the confinement method

### 2013-01-11 Paper/paper.tex
changed paper 11_1

There are however some instances where the method can fail, specially when GDT scores and all the structures are
very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
This protein contains 54 amino acid residues with three alpha helixes and two very short beta strand. 
The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one can find that the 
model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
remain consistent, it may happen that the theoretical model produced by the group "PconsR" much more stable than the 
crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
calculation are within RMSD values of 1.6 - 2.0 \AA. In order to understand the reason behind the mismatch of GDT\_TS values
of crystal and the model1, we have decomposed the total free energy into the per residue component. A close analysis of 
the per residue free energy value, we found that although, crystal and different model have closed GDT\_TS and backbone RMSD 
value, there are number of sidechains oriented in a very different way. As shown in Figure~\ref{fig:T0538compare}(A),
there is a salt bridge between Arg-32 and Glu-35 present in the crystallographic structure, which is absent in the
model 1. In model 1, Arg-32 has another salt bridge with Glu-28. Glu-28 in crystal has the salt bridge with Lys-24. In model 1, Lys-24 is
exposed towards the protein surface. This is shown in Figure~\ref{fig:T0538compare}(A). There are more such interesting and may be crucial
sidechain differences between the model and the crystallographic structure, which indicate the limitation of the GDT\_TS based 
ranking of the protein structure. As, shown in Figure~\ref{fig:T0538compare}(B), Arg-26 and Glu-50 has a salt bridge in the crystallographic 
structure. Also, Phe-51 is oriented towards the protein core in the crystallographic structure. In model 1, the salt bridge between Arg-26 and Glu-50
is absent and the Phe-51 is oriented towards the protein surface.

@@

\bibitem{MacCallum2011}
MacCallum, J.; Pérez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
in CASP9. Proteins, 2011, 79, 74-90.

>>>>

There are however some instances where the method can fail, specially when GDT scores and all the structures are
very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
This protein contains 54 amino acid residues with three alpha helixes and two very short beta strand. 
The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one can find that the 
model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
remain consistent, it may happen that the theoretical model produced by the group "PconsR" much more stable than the 
crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
calculation are within RMSD values of 1.6 - 2.0 \AA. In order to understand the reason behind the mismatch of GDT\_TS values
of crystal and the model1, we have decomposed the total free energy into the per residue component. A close analysis of 
the per residue free energy value i(data not shown) reveal that there are number of sidechains oriented in a very different way in crystal and different 
model. This may account for differences in free energy values, inspite of close RMSD and GST\_TS values. As shown in Figure~\ref{fig:T0538compare}(A),
there is a salt bridge between Arg-32 and Glu-35 present in the crystallographic structure, which is absent in the
model 1. In model 1, Arg-32 has another salt bridge with Glu-28. Glu-28 in crystal has the salt bridge with Lys-24. In model 1, Lys-24 is
exposed towards the protein surface. This is shown in Figure~\ref{fig:T0538compare}(A). There are more such interesting and may be crucial
sidechain differences between the model and the crystallographic structure, which indicate the limitation of the GDT\_TS based 
ranking of the protein structure. As, shown in Figure~\ref{fig:T0538compare}(B), Arg-26 and Glu-50 has a salt bridge in the crystallographic 
structure. Also, Phe-51 is oriented towards the protein core in the crystallographic structure. In model 1, the salt bridge between Arg-26 and Glu-50
is absent and the Phe-51 is oriented towards the protein surface.

@@

\bibitem{MacCallum2011}
MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
in CASP9. Proteins, 2011, 79, 74-90.
